Summary of the things that I learned

### 2023-09-01 transcript/summary.tex
Complete the definition of the functor from objects in a category to algebras for the free functor algebraic theory
Add the last couple of weeks to the summary of learned things, up to the construction of a lambda-theory from an algebra

\usepackage{hyperref}
\usepackage{tikz-cd}

\newcommand{\Catc}[1]{\mathcal{#1}}
\newcommand{\BB}{\Catc{B}}
\newcommand{\CC}{\Catc{C}}
\newcommand{\DD}{\Catc{D}}
\newcommand{\EE}{\Catc{E}}
\newcommand{\Catb}[1]{\mathbf{#1}}
\newcommand{\SET}{\Catb{Set}}
\newcommand{\TYPE}{\Catb{Type}}
\newcommand{\FIN}{\Catb{Fin}}
\newcommand{\Ob}[1]{{#1}_0}
\newcommand{\Hom}[3]{{#1}(#2,#3)}
\newcommand{\op}[1]{\ensuremath{{#1}^\text{op}}}

@@

@@

  \section{Things that I would like to have a standard for}
  \begin{itemize}
    \item Naming (of folders, files, terms): what abbreviations are allowed? In what order do words appear? How are words separated? How are names capitalized?
    \item Documentation headers in files.
    \item In which cases do we want to use unicode?
    \item In which cases do we want definitions to be (fully) typed?
    \item In which cases do we want accessor functions for all data and properties?
  \end{itemize}

@@

  \chapter{Week 08}
  \section{Univalent Categories}
  A univalent category is a category in which the univalence axiom holds. I.e., a category $ \mathcal C $ in which, for all $ A, B \in \Ob \CC $, the canonical map $ (A =_{\mathcal C} B) \to (A \cong B) $ is an equivalence.

@@

  \chapter{Week 09}

@@

  \chapter{Week 11}

@@

  \chapter{Week 17}
  \section{The applicability of algebraic theories}
  It turns out that it is very well possible to characterize a family of algebraic constructs using algebraic theories. For example, if we take the algebraic theory $ \mathcal T $ with $ \mathcal T(n) $ the monoids on $ n $ generators, there is an equivalence (at least of types, maybe also of categories) between $ \mathcal T $-algebras and monoids. We can do the same for groups, rings, ring modules, ring algebras etc. The action of certain elements corresponds then to group operation, the ring multiplication, the ring action, etc:
  \[ a \cdot b := \alpha_2(x_1 \cdot x_2, (a, b)). \]
  The laws that these operations must obey can also be deduced from the action, by pulling back the equation to the free object. For example, associativity:
  \begin{align*}
    a \cdot (b \cdot c) &= \alpha_2(x_1 \cdot x_2, (a, \alpha_2(x_1 \cdot x_2, (b, c))))\\
    &= \alpha_2(x_1 \cdot x_2, (\alpha_3(x_1, (a, b, c)), \alpha_2(x_1 \cdot x_2, (\alpha_3(x_2, (a, b, c)), \alpha_3(x_3, (a, b, c))))))\\
    &= \alpha_3(x_1 \cdot (x_2 \cdot x_3), (a, b, c))\\
    &\dots\\
    &= (a \cdot b) \cdot c.
  \end{align*}
  However, not all algebraic objects can be expressed as algebras for some algebraic theory in this way. Consider, for example, fields. If we have some algebraic theory $ \mathcal T $ that has the fields as its algebras, we would expect some element $ x_1^{-1} = \frac{1}{x_1} \in \mathcal T(1) $ that expresses the multiplicative inverse. However, there is no good way to give a value to $ \alpha_1(x_1^{-1}, (0)) $, since $ 0 $ does not have a multiplicative inverse.

  In general, given a category $ \CC $, a functor $ F : \CC \to \SET $ with a left adjoint $ G : \SET \to \CC $. Then we can set
  \[ T(n) = F(G(\{1, \dots, n\})). \]
  Given $ c \in \Ob{\CC} $, we have a bijection $ \varphi: \Hom{\SET}{\{1, \dots, n\}}{F(c)} \xrightarrow{\sim} \Hom{\CC}{G(\{1, \dots, n\})}{c} $. Then, for $ f \in T(m) $, $ g: \{ 1, \dots, m \} \to T(n) $, we can set $ f \bullet g := F(\varphi(g))(f) $. We also take $ \pi_{n, i} = \varphi^{-1}(\mathrm{id}_{G(\{1, \dots, n\})})(i) $. Then, given any $ A \in \Ob \CC $, we can make $ F(A) $ into an algebra by setting $ \alpha_n(f, a) = F(\varphi(a))(f) $. Also, given a morphism $ f \in \Hom{\CC}{A}{B} $, we have a function $ F(f) \in \Hom{\SET}{F(A)}{F(B)} $ (\textbf{TODO: Show that it is an algebra morphism}). Of course, $ F $ respects composition and identities, so we have a functor from objects of $ \CC $ to $ T $-algebras.

@@

>>>>

\usepackage{hyperref}
\usepackage{tikz-cd}
\usepackage{amssymb}

\newcommand{\Catc}[1]{\mathcal{#1}}
\newcommand{\BB}{\Catc{B}}
\newcommand{\CC}{\Catc{C}}
\newcommand{\DD}{\Catc{D}}
\newcommand{\EE}{\Catc{E}}
\newcommand{\Catb}[1]{\mathbf{#1}}
\newcommand{\SET}{\Catb{Set}}
\newcommand{\TYPE}{\Catb{Type}}
\newcommand{\FIN}{\Catb{Fin}}
\newcommand{\Ob}[1]{{#1}_0}
\newcommand{\Hom}[3]{{#1}\left(#2,#3\right)}
\newcommand{\op}[1]{\ensuremath{{#1}^\text{op}}}

\newtheorem{lemma}{Lemma}

\theoremstyle{definition}
\newtheorem{remark}{Remark}

@@

  \texttt{transportf\_set}

@@

  The pushout of \texttt{exact}, \texttt{apply}, \texttt{refine} and \texttt{use}.

  \section{Things that I would like to have a standard for}
  \begin{itemize}
    \item Naming (of folders, files, terms): what abbreviations are allowed? In what order do words appear? How are words separated? How are names capitalized?
    \item Documentation headers in files.
    \item In which cases do we want to use unicode?
    \item In which cases do we want definitions to be (fully) typed?
    \item In which cases do we want accessor functions for all data and properties?
    \item What should be documented using comments, and how should they be formatted?
  \end{itemize}

@@

  \chapter*{Week 08}
  \section{Univalent Categories}
  A univalent category is a category in which the univalence axiom holds. I.e., a category $ \mathcal C $ in which, for all $ A, B \in \Ob \CC $, the canonical map $ (A =_{\mathcal C} B) \to (A \cong B) $ is an equivalence.

@@

  \chapter*{Week 09}

@@

  \chapter*{Week 11}

@@

  \chapter*{Week 17}
  \section{The applicability of algebraic theories}
  It turns out that it is very well possible to characterize a family of algebraic constructs using algebraic theories. For example, if we take the algebraic theory $ \mathcal T $ with $ \mathcal T(n) $ the monoids on $ n $ generators, there is an equivalence (at least of types, maybe also of categories) between $ \mathcal T $-algebras and monoids. We can do the same for groups, rings, ring modules, ring algebras etc. The action of certain elements corresponds then to group operation, the ring multiplication, the ring action, etc:
  \[ a \cdot b := \alpha_2(x_1 \cdot x_2, (a, b)). \]
  The laws that these operations must obey can also be deduced from the action, by pulling back the equation to the free object. For example, associativity:
  \begin{align*}
    a \cdot (b \cdot c) &= \alpha_2(x_1 \cdot x_2, (a, \alpha_2(x_1 \cdot x_2, (b, c))))\\
    &= \alpha_2(x_1 \cdot x_2, (\alpha_3(x_1, (a, b, c)), \alpha_2(x_1 \cdot x_2, (\alpha_3(x_2, (a, b, c)), \alpha_3(x_3, (a, b, c))))))\\
    &= \alpha_3(x_1 \cdot (x_2 \cdot x_3), (a, b, c))\\
    &\dots\\
    &= (a \cdot b) \cdot c.
  \end{align*}
  However, not all algebraic objects can be expressed as algebras for some algebraic theory in this way. Consider, for example, fields. If we have some algebraic theory $ \mathcal T $ that has the fields as its algebras, we would expect some element $ x_1^{-1} = \frac{1}{x_1} \in \mathcal T(1) $ that expresses the multiplicative inverse. However, there is no good way to give a value to $ \alpha_1(x_1^{-1}, (0)) $, since $ 0 $ does not have a multiplicative inverse.

  In general, given a category $ \CC $, a functor $ F : \CC \to \SET $ with a left adjoint $ G : \SET \to \CC $. Then we can set
  \[ T(n) = F(G(\{1, \dots, n\})). \]
  Given $ c \in \Ob{\CC} $, we have a natural bijection $ \varphi: \Hom{\SET}{\{1, \dots, n\}}{F(c)} \xrightarrow{\sim} \Hom{\CC}{G(\{1, \dots, n\})}{c} $. Then, for $ f \in T(m) $, $ g: \{ 1, \dots, m \} \to T(n) $, we can set $ f \bullet g := F(\varphi(g))(f) $. We also take $ \pi_{n, i} = \varphi^{-1}(\mathrm{id}_{G(\{1, \dots, n\})})(i) $. Then, given any $ A \in \Ob \CC $, we can make $ F(A) $ into an algebra by setting $ \alpha_n(f, a) = F(\varphi(a))(f) $. Also, given a morphism $ m \in \Hom{\CC}{A}{B} $, we have a function $ F(m) \in \Hom{\SET}{F(A)}{F(B)} $.
  are equal. Naturality gives for all $ a: \{1, \dots, n\} \to F(A) $ that
  \[ m \circ \varphi(a) = \varphi(F(m) \circ a), \]
  so for $ f: F(G(\{1, \dots, n\})) $,
  \[ F(m)(\alpha_2(f, a)) = F(m \circ \varphi(a))(f) = F(\varphi(F(m) \circ a))(f) = \alpha_2(f, F(m) \circ a) \]
  and $ F(m) $ is a morphism of algebras. Of course, $ F $ respects composition and identities, so we have a functor from objects of $ \CC $ to $ T $-algebras.

@@

  \chapter*{Week 18}

  \begin{lemma}
    The lambda calculus $ \Lambda $ is the initial $ \lambda $-theory.
  \end{lemma}
  \begin{proof}
    Given a $ \lambda $-theory $ \mathcal L $, we need a unique morphism $ \varphi : \Lambda \to \mathcal L $.

    Do structural induction on the lambda terms in $ \Lambda(n) $. Send \texttt{Var(i)} to $ \pi_{n, i} $, \texttt{Abs(f)} to $ \lambda(\varphi(f)) $ and \texttt{App(f, g)} to $ \rho(\varphi(f)) \bullet (\pi_1, \dots, \pi_n, g) $.

    Show that it preserves $ \bullet $, $ \pi_{i} $, $ \rho $ and $ \lambda $.

    Uniqueness: Again, structural induction on the lambda terms. Preservation of the $ \pi_i $ gives the image of \texttt{Var(i)}. Preservation of $ \lambda $ gives the image of \texttt{Abs(f)}. Preservation of $ \rho $ and $ \bullet $ gives the image of
    \[ \mathtt{App(f, g)} = \rho(f) \bullet (\pi_1, \dots, \pi_n, g). \]
  \end{proof}

  Given a $ \lambda $-theory $ \mathcal L $, $ \mathcal L(0) $ is a $ \Lambda $-algebra (by extension of scalars). This gives a functor.

  \begin{lemma}
    Given a $ \Lambda $-algebra $ A $, we can form the set $ \mathcal U_A $.
  \end{lemma}
  \begin{proof}

  \end{proof}
  \begin{lemma}
    $ \mathcal U_A $ is a $ \lambda $-theory.
  \end{lemma}

  \begin{lemma}
    $ \mathcal U_A $ is functorial in $ A $.
  \end{lemma}
  \begin{proof}

  \end{proof}

  These functors form an adjoint equivalence: We have natural isomorphisms $ \mathcal U_{\mathcal L(0)} \cong \mathcal L $ and $ \mathcal U_A(0) \cong A $.

  \chapter*{Week 33}
  \section{Coends}
  In a category $ \DD $ with equalizers and products, and given a functor $ F : \op \CC \times \CC \to \DD $, we can define the end $ \int_{c: \Ob \CC} F(c, c) $ to be the equalizer of
  \[ \prod_{c : \Ob \CC} F(c, c) \rightrightarrows \prod_{\Hom{\CC} c {c^\prime}} F(c, c^\prime). \]
  For $ \DD = \SET $, an element of this gives, for every element $ c : \Ob \CC $, an element $ f_c : F(c, c) $, such that for all morphisms $ \alpha : \Hom{\CC}{c}{c^\prime} $, we have
  \[ \#F(c, \alpha)(f_c) =_{F(c, c^\prime)} \#F(\alpha, c^\prime)(f_{c^\prime}). \]

  If we have two functors $ G, G^\prime: \CC \to \CC^\prime $, take $ F(c, c^\prime) = \Hom{\DD^\prime}{G(c)}{G^\prime(c^\prime)} $. An element of $ \int_c \Hom{\DD^\prime}{G(c)}{G^\prime(c^\prime)} $ gives, for all $ c: \Ob \CC $ an element $ f_c : \Hom{\SET}{G(c)}{G(c^\prime)} $ such that for all morphisms $ \alpha : \Hom{\CC}{c}{c^\prime} $, we have
  \[ f_c \cdotp \# G^\prime(\alpha) = \#F(c, \alpha)(f_c) = \#F(\alpha, c^\prime)(f_{c^\prime}) = \#G(\alpha) \cdotp f_{c^\prime}, \]
  so we have a commutative diagram
  \begin{center}
    \begin{tikzcd}
      G(c) \arrow[r, "\#G(\alpha)"] \arrow[d, "f_c"] & G(c^\prime) \arrow[d, "f_{c^\prime}"]\\
      G^\prime(c) \arrow[r, "\#G^\prime(\alpha)"] & G^\prime(c)
    \end{tikzcd}
  \end{center}
  Which is exactly the diagram of a natural transformation.

  On the other hand, if $ \DD $ has coproducts, we can define the coend $ \int^{c: \Ob \CC} F(c, c) $ to be the coequalizer of
  \[ \coprod_{\Hom{\CC} c {c^\prime}} F(c, c^\prime) \rightrightarrows \coprod_{c : \Ob \CC} F(c, c). \]
  For $ \DD = \SET $, an element of this is an element $ c : \Ob \CC $, together with an element of $ F(c, c) $.
  For $ c, c^\prime : \Ob \CC $, $ f : F(c^\prime, c) $ and $ \alpha : \Hom{\CC}{c}{c^\prime} $, we identify
  \[ (c, \#F(\alpha, c)(f)) \]
  with
  \[ (c^\prime, \#F(c^\prime, \alpha)(f)). \]

  For $ F(c, c^\prime) = \Hom{\CC^\prime}{G(c)}{G^\prime(c)} $, an element of $ \int^{c: \Ob \CC} $ is an element $ c : \Ob \CC $ with a morphism $ f : \Hom{\CC^\prime}{G(c)}{G^\prime(c)} $ and for a morphism $ \alpha : \Hom{\CC}{c}{c^\prime} $ and morphism $ f : \Hom{\CC^\prime}{G(c^\prime)}{G^\prime(c)} $, we set
  \[ (c, \#G(\alpha) \cdotp f) = (c, \#F(\alpha, c)(f)) = (c^\prime, \#F(c^\prime, \alpha)(f)) = (c^\prime, f \cdotp \#G^\prime(\alpha)). \]

  \subsection{Wedges}
  A wedge $ e : w \to F $ is an object $ w $ with, for every object $ c : \Ob \CC $ a morphism $ e_c : w \to F(c, c) $ and for every morphism $ \alpha: c \to c^\prime $ a commutative square
  \begin{center}
    \begin{tikzcd}
      w \arrow[d, "e_c"] \arrow[r, "e_{c^\prime}"] & {F(c^\prime, c^\prime)} \arrow[d, "{\#F(\alpha, c^\prime)}"] \\
      {F(c, c)} \arrow[r, "{\#F(c, \alpha)}"] & {F(c, c^\prime)}
    \end{tikzcd}
  \end{center}
  The end $ e : \int_{c: \Ob \CC} F(c, c) \to F $ is the universal wedge. I.e., given a wedge $ e^\prime : w^\prime \to F $, $ e^\prime $ factors through $ e $ via a unique map $ w^\prime \to \int_{c: \Ob \CC} F(c, c) $.

  Conversely, a cowedge $ e : F \to w $ is an object $ w $ with, for every object $ c : \Ob \CC $ a morphism $ e_c : F(c, c) \to w $ and for every morphism $ \alpha: c \to c^\prime $ a commutative square
  \begin{center}
    \begin{tikzcd}
      {F(c^\prime, c)} \arrow[r, "{\#F(c^\prime, \alpha)}"] \arrow[d, "{\#F(\alpha, c)}"] & {F(c^\prime, c^\prime)} \arrow[d, "e_{c^\prime}"] \\
      {F(c, c)} \arrow[r, "e_c"] & w
    \end{tikzcd}
  \end{center}
  The coend $ e : F \to \int^{c: \Ob \CC} F(c, c) $ is the universal cowedge. I.e., given a cowedge $ e^\prime : F \to w^\prime $, $ e^\prime $ factors through $ e $ via a unique map $ \int_{c: \Ob \CC} F(c, c) \to w^\prime $.

  \subsection{The coproduct of algebras}
  Now, to the example that we are interested in
  \[ A + B = \int^{m,n : \Ob F} \mathrm{Alg}_T(T(m), A) \times \mathrm{Alg}_T(T(n), B) \times T(m + n). \]
  We can identify $ \mathrm{Alg}_T(T(m), A) \cong A^m $:
  \begin{itemize}
    \item Send $ f : \mathrm{Alg}_T(T(m), A) $ to $ (f \pi_1, \dots, f \pi_m) : A^m $;
    \item Send $ (a_1, \dots, a_m) : A^m $ to $ x \mapsto x \bullet (a_1, \dots, a_m) : A $.
  \end{itemize}
  Then this becomes
  \[ A + B = \int^{m,n : \Ob F} A^m \times B^n \times T(m + n). \]
  An element of this set is a pair
  \[ ((m, n), ((a_1, \dots, a_m), (b_1, \dots, b_n), f)) \]
  with $ m, n : \Ob F $, $ a_i : A $, $ b_i : B $ and $ f : T(m + n) $.

  If we have $ m, m^\prime, n, n^\prime : \Ob F $, $ \alpha : \Hom{F}{m}{m^\prime} $, $ \beta : \Hom{F}{n}{n^\prime} $, and an element
  \[ ((a_i)_i, (b_i)_i, f) : A^{m^\prime} \times B^{n^\prime} \times T(m + n), \]
  we identify
  \[ ((m, n), ((a_{\alpha(i)})_i, (b_{\beta(i)})_i, f)) = ((m^\prime, n^\prime), ((a_i)_i, (b_i)_i, \#T(\alpha + \beta)(f))) \]

  Now, we can define the action
  \[ f \bullet (((m, m^\prime), ((a_{ji})_i, (b_{ji})_i, g_j))_j) := ((m, m^\prime), (f \bullet (a_{ji})_j)_i, (f \bullet (b_{ji})_j)_i, f \bullet (g_j)_j). \]
  This is associative and projects components because the action on $ A $ and $ B $ is associative and projects components.

  \subsubsection{The universal property}
  Now, we would just need to show that for all $ C : \Ob{\mathrm{Alg}_T} $ and all $ f_A : A \to C $ and $ f_B : B \to C $, there exists a unique morphism $ g : A + B \to C $.

  Since $ A + B $ is a coend, it would suffice to give $ C $ a $ A^m \times B^n \times T(m + n) $-cowedge structure. That is, we need, for all $ m, n: \Ob F $, a morphism $ A^m \times B^n \times T(m + n) \to C $. Define
  \[ f_{A + B}((a_1, \dots, a_m), (b_1, \dots, b_n), f) := f \bullet (f_A(a_1), \dots, f_A(a_m), f_B(b_1), \dots, f_B(b_n)). \]
  We must show that given $ x: T(l) $ and given, for all $ 1 \leq i \leq l $, $ (a_i, b_i, f_i) : A^m \times B^n \times T(m + n) $, we have
  \[ x \bullet (f_{A + B}(a_i, b_i, f_i))_i = f_{A + B}((x \bullet (a_i, b_i, f_i))_i). \]
  But
  \begin{align*}
    &x \bullet (f_i \bullet ((f_A(a_{ij}))_j, (f_B(b_{ij}))_j))_i\\
    &=x \bullet (f_{A + B}(a_i, b_i, f_i))_i\\
    &= f_{A + B}((x \bullet (a_i, b_i, f_i))_i)\\
    &= f_{A + B}((x \bullet (a_{ij})_i)_j, (x \bullet (b_{ij})_i)_j, x \bullet f)\\
    &= (x \bullet f) \bullet ((f_A(x \bullet (a_{ij})_i))_j, (f_B(x \bullet (b_{ij})_i))_j)\\
    &= (x \bullet f) \bullet ((x \bullet (f_A(a_{ij}))_i)_j, (x \bullet (f_B(b_{ij}))_i)_j)\\
  \end{align*}

  \chapter*{Week 34}
  \section{Algebraic Theories as categories}
  Hyland cites \cite{p13-adamek-rosicky}, which also introduces algebraic theories. There is a correspondence between Hyland's algebraic theories, and ``one-sorted algebraic theories'' in the book.

  The book defines the category $ \mathcal N $, which is the full subcategory of $ \op \SET $, with objects $ n = \{ 0, 1, \dots, n - 1 \} $ for all $ n \in \mathbb N $. This category presents $ n $ as $ 1^n $, with projections $ \pi_{n, i} : n \to 1 $.

  \subsection{Algebraic Theories}
  The book defines a one-sorted algebraic theory to be a category with finite products, whose objects are the natural numbers, together with a functor $ T : \mathcal N \to \mathcal T $ that is the identity on objects.

  Given a one-sorted algebraic theory $ (\mathcal T, T) $, we make Hyland's algebraic theory $ \mathcal T^\prime $ as follows:
  Take $ \mathcal T^\prime(n) = \mathcal T(n, 1) $. We have $ \mathcal T^\prime(n)^m = \mathcal T(n, 1)^m = \mathcal T(n, m) $. Then composition gives
  \[ \bullet : \mathcal T^\prime(m) \times \mathcal T^\prime(n)^m = \mathcal T(m, 1) \times \mathcal T(n, m) \to \mathcal T(n, 1) = \mathcal T^\prime(n). \]
  The projections are given by $ \pi^\prime_{n, i} = \#T(\pi_{n, i}) : \mathcal T(n, 1) = \mathcal T^\prime(n) $.

  Conversely, given Hyland's algebraic theory $ \mathcal T^\prime $, we make a one-sorted algebraic theory $ \mathcal T $ as follows. Take $ \Ob{\mathcal T} = \mathbb N $ and $ \Hom{\mathcal T}{m}{n} = \mathcal T^\prime(m)^n $. Then $ T $ sends $ f = (\pi_{i_1}, \dots, \pi_{i_n}) = \mathcal N(m, 1)^n = \mathcal N(m, n) $ to $ (\pi^\prime_{i_1}, \dots, \pi^\prime_{i_n}) $. Composition is then defined as
  \[ \mathcal T(l, m) \times \mathcal T(m, 1)^n = \mathcal T(l, m) \times \mathcal T(m, n) \ni (f, (g_1, \dots, g_n)) = (f, g) \mapsto (g_1 \bullet f, \dots, g_n \bullet f) : \mathcal T(l, 1)^n = \mathcal T(l, n). \]

  \subsection{Algebraic theory algebras}
  Given a one-sorted algebraic theory $ (\mathcal T, T) $ and the corresponding Hyland's algebraic theory $ \mathcal T^\prime $.

  An algebra for $ \mathcal T $ is a finite-product-preserving functor $ A: \mathcal T \to \SET $. Given such an algebra, we get an algebra $ A^\prime $ for $ \mathcal T^\prime $ by setting $ A^\prime = A(1) $, and for $ f : \mathcal T^\prime(n) = \mathcal T(n, 1) $, we define $ f \bullet - $ to be
  \[ \#A(f) : (A^\prime)^n = A(1)^n = A(1^n) = A(n) \to A(1) = A^\prime. \]
  Conversely, given an algebra $ A^\prime $ for $ \mathcal T^\prime $, we define the functor $ A $ to map $ n $ to $ (A^\prime)^n $ on objects, and to send the morphism $ f : \mathcal T(m, n) = \mathcal T(m, 1)^n = (\mathcal T^\prime(m))^n $ to
  \[ A(m) = (A^\prime)^m \ni a \mapsto (f_1 \bullet a, \dots, f_n \bullet a) (A^\prime)^n = A(n). \]
  \subsection{Pre(- or post)liminaries}
  \subsubsection{The Density Theorem}
  Given an algebraic theory (can be one-sorted) $ \mathcal T $, we have a Yoneda embedding $ Y_{\mathcal T}: \op{\mathcal T} \to [\mathcal T, \SET] $, sending $ X $ to $ \Hom{\mathcal T}{X}{-} $. Also, given a $ \mathcal T $-algebra $ A $, we have the category of elements $ \mathop{El} A $, with the objects being dependent pairs $ \sum_{x : \Ob{\mathcal T}}, A x $ and morphisms $ \Hom{\mathrm{El}A}{(X, x)}{(Y, y)} $ being morphisms $ f : \Hom{\mathcal T}{Y}{X} $ with $ \#A(f)(y) = x $. Denote $ \Phi_A : \mathop{El} A \to \op{\mathcal T} $ the projection on the first coordinate. Then $ A $ is the colimit of the diagram
  \[ \mathop{El} A \xrightarrow{\Phi_A} \op{\mathcal T} \xrightarrow{Y_{\mathcal T}} [\mathcal T, \SET]. \]
  I.e. we have, for all $ (x, y) : \mathop{El} A $, a map of functors (i.e. natural transformation)
  \[ Y_{\mathcal T}(\Phi_A(x, y)) = Y_{\mathcal T}(x) = \Hom{\mathcal T}{x}{-} \Rightarrow A \]
  That is, for all $ t : \mathcal T $, a map
  \[ \mathcal T(x, t) \to A(t), f \mapsto \#A(f)(y) \]
  and for all morphisms $ g : \Hom{\mathcal T}{t}{t^\prime} $ and all $ f: \mathcal T(x, t) $, we have the equality
  \[ \#A(g)(\#A(f)(y)) = \#A(g \circ f)(y) \]
  so the map is natural.

  Now, for all $ B : [\mathcal T, \SET] $, we have
  \begin{align*}
    \Hom{[\mathcal T, \SET]}{\mathop{colim}_{(x, y) : \mathop{El} A} Y_{\mathcal T}(\Phi_A(x, y))}{B} &\cong \lim_{(x, y) : \mathop{El} A} \Hom{[\mathcal T, \SET]}{Y_{\mathcal T}(\Phi_A(x, y))}{B}\\
    &\cong \lim_{(x, y) : \mathop{El} A} \Hom{[\mathcal T, \SET]}{Y_{\mathcal T}(x)}{B}\\
    &\cong \lim_{(x, y) : \mathop{El} A} B(x)\\
    &\cong \Hom{[\mathop{El} A, \SET]}{\mathrm{pt}}{B}
  \end{align*}
  for $ \mathrm{pt} $ the constant functor that sends anything to $ \{ \star \} $. Now, given a natural transformation $ \alpha : \Hom{[\mathop{El} A, \SET]}{\mathrm{pt}}{B} $. For all $ x : \Ob{\mathcal T} $, we get a map
  \[ A(x) \to B(x), \quad y \mapsto \alpha_{(x, y)}(\star). \]
  By naturality of $ \alpha $, this is a natural transformation $ A \Rightarrow B $. This gives an isomorphism
  \[ \Hom{[\mathop{El} A, \SET]}{\mathrm{pt}}{B} \cong \Hom{[\mathcal T, \SET]}{A}{B}. \]
  Since this holds for all $ B $, we have
  \begin{align*}
    Y_{\op{[\mathcal T, \SET]}}(A)
    &= \Hom{[\mathcal T, \SET]}{A}{-}\\
    &\cong \Hom{[\mathcal T, \SET]}{\mathop{colim}_{(x, y) : \mathop{El} A} Y_T(\Phi_A(x, y))}{-}\\
    &= Y_{\op{[\mathcal T, \SET]}}(\mathop{colim}_{(x, y) : \mathop{El} A} Y_T(\Phi_A(x, y))).
  \end{align*}
  By the Yoneda Lemma, this implies that $ \mathop{colim}_{(x, y) : \mathop{El} A} Y_T(\Phi_A(x, y)) \cong A $.

  \subsection{An example}
  First of all, we need an example. Take an algebraic theory (Hyland's kind): $ \mathcal T^\prime(n) = \{\star, 1, 2, 3, \dots, n \} $ with $ \star \bullet g = \star $ and $ i \bullet g = g_i $. Take the two-element set $ A^\prime = \{ \bot, \top \} $. We can make $ A^\prime $ into an $ \mathcal T^\prime $-algebra by setting $ i \bullet a := a_i $ and $ \star \bullet a = \top $ (or $ \bot $ for that matter, as long as we are consistent).

  Translating this to a one-sorted algebraic theory, we get the category $ \mathcal T $ with $ \Ob{\mathcal T} = \{ 0, 1, 2, \dots \} $ and $ \mathcal T(m, n) = \{ \star, 1, \dots, m \}^n $. We also get the algebraic theory (i.e. functor) $ A: \mathcal T \to \SET $, given on objects by $ n \mapsto \{ \bot, \top \}^n $ and on morphisms by
  \[ \mathcal T(m, 1) \ni f \mapsto a \mapsto \left\{ \begin{array}{cc} a_i & f = i\\ \top & f = \star \end{array} \right. \]
  and extending linearly for $ f \in \mathcal T(m, n) = \mathcal T(m, 1)^n $. The category $ \mathrm{El} A $ consists of objects $ (n, a) $ with $ a : \{ \bot, \top \}^n $. Its morphisms $ f: \Hom{\mathrm{El} A}{(m, a)}{(n, b)} $ are morphisms $ \Hom{\mathcal T}{n}{m} = \mathcal T^\prime(n)^m $ such that $ A(f)(b) = (f_1 \bullet b, \dots, f_m \bullet b) = a $, so
  \[ \Hom{\mathrm{El} A}{(m, a)}{(n, b)} \cong \prod_{1 \leq i \leq n} \Hom{\mathrm{El} A}{(1, a_i)}{(n, b)}, \]
  and
  \[ (m, a) \cong \coprod_{1 \leq i \leq n} (1, a_i). \]
  In other words, all objects of $ \mathrm{El} A $ are generated as coproducts by $ (1, \top) $ and $ (1, \bot) $. Now, we have $ \Hom{\mathrm{El} A}{(1, a)}{(n, b)} = \{ f : \mathcal T^\prime(n) \mid f \bullet b = a. \} $. This is the set of $ i $ such that $ b_i = a $, and if $ a = \top $, it contains $ \star $. Note that $ \mathcal T(n, 1) \cong \coprod_{a : A^\prime} \Hom{\mathrm{El} A}{(n, c)}{(1, a)} $.

  Now, for an object $ n : \op{\mathcal T} $, the Yoneda embedding sends it to
  \[ m \mapsto \Hom{\mathcal T}{n}{m} = \mathcal T^\prime(n)^m \]
  A morphism $ f: \Hom{\op{\mathcal T}}{n}{n^\prime} = \Hom{\mathcal T}{n^\prime}{n} $ is sent to a natural transformation from $ m \mapsto \Hom{\mathcal T}{n}{m} $ to $ \Hom{\mathcal T}{n^\prime}{m} $ by sending $ g : \Hom{\mathcal T}{n}{m} \cong \mathcal T^\prime(n)^m $ to
  \[ g \circ f = (g_1 \bullet f, \dots, g_n \bullet f) : \Hom{\mathcal T}{n^\prime}{m} \cong T^\prime(n)^m. \]

  We can calculate (co)limits in a functor category pointwise. That means that
  \[ (\mathop{colim}_{j : J} \mathcal F(j))(x) = \mathop{colim}_{j : J} (\mathcal F(j)(x)) \]
  We are first and foremost interested in
  \begin{align*}
    \left(\mathop{colim}_{(n, a) : \mathrm{El} A} Y_{\mathcal T}(n)\right)(1)
    &= \mathop{colim}_{(n, a) : \mathrm{El} A} (Y_{\mathcal T}(n)(1))\\
    &= \mathop{colim}_{(n, a) : \mathrm{El} A} (\mathcal T(n, 1))\\
    &= \mathop{colim}_{(n, a) : \mathrm{El} A} (\mathcal T^\prime(n))
  \end{align*}
  which is the equalizer of
  \[ \coprod_{f: (m, a) \to (m^\prime, a^\prime)} \mathcal T^\prime(m) \rightrightarrows \coprod_{(n, a) : \mathrm{El} A} \mathcal T^\prime(n) \]
  for every $ f: (m, a) \to (m^\prime, a^\prime) $ (i.e. $ f: \mathcal T^\prime(m^\prime)^m $ such that for all $ i $, $ f_i \bullet a^\prime = a_i $), we identify the $ t : \mathcal T^\prime(m) $ over $ (m, a) $ with the $ t \bullet f : \mathcal T^\prime(m^\prime) $ over $ (m^\prime, a^\prime) $.

  In particular, we identify $ \pi_1 : \mathcal T^\prime(1) $ over $ (1, f \bullet a) $ with $ \pi_1 \bullet f = f : \mathcal T^\prime(m) $ over $ (m, a) $.

  For general $ m $ and $ m^\prime $, we identify $ t \bullet f $ over $ (m^\prime, a) $ with $ t $ over $ (m, (f_1 \bullet a, \dots, f_n \bullet a)) $. Since the action on the algebra is associative, we have
  \[ (t \bullet f) \bullet a = t \bullet (f_1 \bullet a, \dots, f_n \bullet a) \]
  so the identifications of the $ t $ over $ (m, a) $ with $ \pi_1 $ over $ (1, t \bullet a) $ generate the equivalence relation.

  Now, for the action of $ \mathcal T $ on this colimit. In particular, given $ f: \Hom{\mathcal T}{m}{1} = \mathcal T^\prime(m) $, we want a morphism
  \[ \Hom{\SET}{\left(\mathop{colim}_{(n, a) : \mathrm{El} A} Y_{\mathcal T}(n)\right)(m)}{\left(\mathop{colim}_{(n, a) : \mathrm{El} A} Y_{\mathcal T}(n)\right)(1)}. \]
  We get this as
  \[ \mathop{colim}_{(n, a) : \mathrm{El} A} (t \mapsto f \bullet t): \Hom{\SET}{\mathop{colim}_{(n, a) : \mathrm{El} A} (\mathcal T^\prime(n)^m)}{\mathop{colim}_{(n, a) : \mathrm{El} A} (\mathcal T^\prime(n))}. \]
  Every element can be identified with $ \pi_1 $ over $ (1, a) $ for some $ a $. Now, for $ a_1, \dots, a_n: A^\prime $, $ \pi_1 $ over $ (1, a_i) $ is identified with $ \pi_i $ over $ (n, (a_1, \dots, a_n)) $. Then the action of $ f $ sends $ (\pi_1, \dots, \pi_n) $ over $ (n, a) $ to $ f \bullet (\pi_1, \dots, \pi_n) = f $ over $ (n, a) $, but this is identified with $ \pi_1 $ over $ (1, f \bullet a) $. Therefore, this recovers the action of $ \mathcal T^\prime $ on $ A^\prime $.

  \chapter*{Week 35}
  \section{The monoid}
  \textit{Note to self: even though everything is defined by actions of $ \Lambda $, it might be good to define application and abstraction in $ A $.}
